Tabular Constraint Learning

### 2016-04-11 Paper/ECAI/tabular.tex

\subsection{Groups, Type-consistency and Constraints}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called \textit{type-consistent} iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. \textit{\CName} is the textual name of the constraint together with its variable names. \textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g. the length of vectors in the arguments must be equal. These are constraints on the group meta-information not on the actual group content. \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function. For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.

\section{Problem Statement}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of Excel constraints and group assignmentsd as follows:

@@

\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies. Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$. If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.

@@

\begin{table*}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    ALLDIFFERENT(X) & & \\
    X = Y & & \\
    FK $\rightarrow$ PK & & \\
    R = PRODUCT(FV, FK=PK | PV) & & \\
    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
    FV = LOOKUP(FK, PK, PV) & & \\
    PERMUTATION(X) & & \\
    R = O1 * O2 & & \\
    R = PROJECT(P) & & \\
     Y = RANK(X)   & & \\
    A = PREV(A) + P - N & & \\
    SERIES(X) & & \\
    Y = SUM(X, col) & & \\
    Y = SUM(X, row) & & \\
    R = SUMIF(FK=PK, V) & & \\
    R = SUMPRODUCT(O1, O2) & & \\

@@

\subsection{Workflow}
\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$T$ is \textbf{not} provided}
      \State $T \gets \extracttables(D)$
    \EndIf
    \If{$G$ is \textbf{not} provided}
      \State $G \gets \extractgroups(D, T)$
    \EndIf
    \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
    \State $S \gets \postprocess(S)$
    \State \Return $S$
\end{algorithmic}
\caption{Workflow}
\label{algo:workflow}
\end{algorithm}

@@

\section{Declarative modeling}
\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}

@@

{\bfseries 
  Experimental questions
}
\begin{itemize}
  \item  How accurate are we? (Accuracy / recall)
  \item  How fast are we and which factors affect the runtime (how)?
  \item  How general is our approach, what limitations are there?
\end{itemize}

>>>>

\samuel{Explain concepts on example}
\subsection{Groups, Type-consistency and Constraints}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called \textit{type-consistent} iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. \textit{\CName} is the textual name of the constraint together with its variable names. \textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g. the length of vectors in the arguments must be equal. These are constraints on the group meta-information not on the actual group content. \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function. For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
\samuel{2nd problem statement (from approach), remove EQUAL}
\section{Problem Statement}
In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}. Here we formalize the statement in terms of Excel constraints and group assignmentsd as follows:

@@

\samuel{Running example (RANK?)}
\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies. Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$. If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.

@@

\begin{table*}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    ALLDIFFERENT(X) & & \\
    X = Y & & \\
    FK $\rightarrow$ PK & & \\
    % TODO check out
    R = LOOKUP-PRODUCT(FV, FK=PK, PV) & & \\
    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
    FV = LOOKUP(FK, PK, PV) & & \\
    PERMUTATION(X) & & \\
    R = O1 * O2 & & \\
    R = PROJECT(P) & & \\
     Y = RANK(X)   & & \\
    A = PREV(A) + P - N & & \\
    SERIES(X) & & \\
    Y = SUM(X, col) & & \\
    Y = SUM(X, row) & & \\
    R = SUMIF(FK=PK, V) & & \\
    R = SUMPRODUCT(O1, O2) & & \\

\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}

@@

\samuel{Move to beginning without algo}
\subsection{Workflow}
\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$T$ is \textbf{not} provided}
      \State $T \gets \extracttables(D)$
    \EndIf
    \If{$G$ is \textbf{not} provided}
      \State $G \gets \extractgroups(D, T)$
    \EndIf
    \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
    \State \Return $S$
\end{algorithmic}
\caption{Workflow}
\label{algo:workflow}
\end{algorithm}

@@

\samuel{Move next to subgroup satisfaction search}
\section{Declarative modeling}
\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}

@@

{\bfseries 
  Experimental questions
}

\begin{itemize}
  \item  How accurate are we? (Accuracy / recall)
  \item  How fast are we and which factors affect the runtime (how)?
  \item  How general is our approach, what limitations are there?
\end{itemize}
